%Chapter 5 - Evaluating the MEGA65 project against known risks
%

\chapter{Evaluating the MEGA65 project against known risks}
\label{Chapter5}
This chapter aims to provide useful advice to the MEGA65 project with the aim of reducing the project's risk. To achieve this, the detailed snapshot of the project's progress at the point in time of July 2018, chapter \ref{Chapter4}, is evaluated to determine its risk levels in the known risks identified in the case studies in chapter \ref{Chapter3}. The layout of this chapter is that each risk identified in the risk list is given a section and within each section the MEGA65 project is scrutinised to determine it risk level to that specific risk. At the end of the chapter a conclusion is drawn on the MEGA65 project's risk level as of July 2019. For the identified high risk areas, some advice that could reduce the project's risk level is given. This advice is in the form of alternate solutions or methods of doing things and their potential risk reduction. The risks to the MEGA65 project are determined by following the guidelines given in \textit{AS ISO 31000:2018 Risk management — Guidelines} 
\cite{RN164} and categorised using the definition of damage listed in the table at the start of chapter \ref{Chapter3}.
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Law and Regulations}
MEGA65 

ask paul and detlef about laws and regulations

from case study:
- electronic interference emmissions (ECC?)
- suitable age testing (PEGI)

\subsection{Brexit-like events}
There are no foreseeable Brexit-like events that would be likely to affect the MEGA65 project. Brexit itself is unlikely to affect the MEGA65 project as the project is largely based in Australia, with the desktop version being made in Germany, which is a European country. 

Ask Paul about how brexit could effect desktop version.

\section{Project Constraints}
The MEGA65 project is not bound to tight constraints on project cost for the most part. The desktop form factor may be subject to more tight constraints but overall the project was started as a hobby and continues in part as a learning tool for students at Flinders University.

\section{Crowd-funding}
The MEGA65 is planning on using crowd-funding to raise funds but importantly they are waiting until the design and component selection is complete before launching the campaign.

\subsection{Contract  of sale}
The fact the the MEGA65 team are waiting until much later in the productisation process to launch the crowd-funding campaign greatly reduces the risks associated with it. The risk comes from forming a contract of sale between MEGA65 project and backers of the crowd-funding campaign and that contract not being fur filled. This could lead to legal proceedings against the MEGA65 project. The fact that the MEGA65 project won't do this until a much later stage in productisation means the chance that the MEGA65 project will fail to deliver the product is greatly reduced. This is due to the uncertainty factor being reduced as a lot of the design and component selection is done as such is not uncertain but certain.

\subsection{Currency}
Ask paul what currency they plan to crowd fund in


\subsection{Failure to meet goal}
ask paul what they will do if they fail to meet crowd funding goal

\section{Sourcing old components}
MEGA65 plans to use floppy drives, mention this in chapter 4 too!

\section{Use of 3rd party intellectual property}
This is the main problem area:
- MEGA65 wanted to use a Commodore logo on the Commodore keyboard key
ask Paul if they had agreement to use commodore icon
recommend they don't use Commodore icon but a similar design

- MEGA65 NEEDS to use BASIC ROM/s to function as a Commodore compatibly computer
ask paul if they have agreement
recommend they don't use 3rd party BASIC ROM and consider producing their own


\section{Open-source business}
ask paul about business model

\section{Loss of intellectual property}
ask paul who holds rights to everything mega65 related inc. software designed by students, FPGA, firmware and PCB designs. desktop version too, case CAD, keyboard design, PCB

\section{Supplier failure}
zask paul about suppliers, check skype for desktop form factor supplier, make risk assessment on them

\section{Open-source fair use}
should not be a concern for MEGA65

ask paul if they used any open-source components

\section{Physical production problems}
