%Chapter 6 - Re-evaluating the MEGA65 project against known risks
% 	Did the MEGA65 project reduce its risk after 10 months?

\chapter{Risk evaluation of the MEGA65 project in May 2019}
\label{Chapter6}
After a 10 month period, the MEGA65 project was re-evaluated against the risks identified in chapter \ref{Chapter3}. This period of time gave the project sufficient time to act on the suggestions given at the end of chapter \ref{Chapter5}. By re-evaluating the MEGA65 project and comparing the result with the evaluation from July 2019, it is hoped that a conclusion can be drawn on whether the MEGA65 project reduced its risk exposure or not. 
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------

\section{Laws and regulations}
The MEGA65 project has undertaken some research into the laws and regulation which need to be upheld in some of their key markets in which they expect to operate. With the knowledge gained from the research, the uncertainty in the relevant laws and regulations is reduced, as such the overall risk to the project is reduced. Not every market was researched, and the unknown markets still expose the project to risk if the MEGA65 where to be sold within it. The MEGA65 project is rated as having an unlikely chance of occurring and if it causing a marginal level of potential damage. \\

Likelihood - Unlikely \\
Potential damage - Marginal \\
Risk - Low \\


\section{Sourcing old components}
During the interim 10 months the MEGA65 project purchased 700 3.5" floppy disk drives and has them in storage in Germany. These disk drives are to be used as components during the assembly of the MEGA65 desktop version. This action of purchasing a large surplus of stock allows the MEGA65 project to avoid this risk in the medium term. When the stock runs out the risk will increase, as such the risk is largely tied to the MEGA65's popularity and sales volume. If the MEGA65 becomes hugely popular then this risk could increase dramatically. If the MEGA65 sells as expected by the MEGA65 project team, then 700 drives will be sufficient for the medium to long term. The potential damage the MEGA65 project is rated as marginal and the chance of it occurring possible. \\

Likelihood - Possible \\
Potential damage - Marginal \\
Risk - Moderate \\


\section{Use of 3rd party intellectual property}
The MEGA65 project has enacted some of the suggestions from chapter \ref{Chapter5}. The steps undertaken in the identified risk areas are discussed below.

\subsection{Commodore logo}
The MEGA65 team decided to use their own design for the logo on the Commodore keyboard key. They have created and used the design for the pre-production run of the desktop form factor. This new design should completely remove the potential for this risk to effect the MEGA65 project.

\subsection{Commodore 64 character set}
The MEGA65 have decided to replace the Commodore character set with one they have created themselves from other freely available character sets and their own work. By removing the Commodore character set the risk eliminated. 

\subsection{Commodore BASIC ROM and kernel ROM}
The MEGA65 team have decided to created their own BASIC and kernel ROMs. This reimplementation is currently under way and can provide basic functionality at this stage. By removing the Commodore ROMs from the MEGA65, the risk is entirely eliminated. \\

Likelihood - Rare \\
Possible Damage - Negligible \\
Risk - Low \\


\section{Open-source business}
The MEGA65 team have not trademarked the MEGA65 name and are operating under the assumption that the MEGA65 will not be hugely popular. The MEGA65 team have been made aware of the strategies listed in chapter \ref{Chapter5}, these strategies are only sensible if the MEGA65 becomes hugely popular and cannot be enacted to much benefit before. Because the MEGA65 team is aware of these strategies the MEGA65 project's potential damage and chance of occurring are both lowered compared to July 2018. \\

Likelihood - Unlikely \\
Possible Damage - Moderate \\  
Risk - Low \\

\section{Comparing risk evaluations from July 18 with May 19}
By comparing the risk evaluation of the MEGA65 project from July 2018 with the evaluation from May 2019 it can be seen how the MEGA65 project's risk exposure changed during that time. Table \ref{tab:table1} shows the risk evaluation results for the risk areas in which the MEGA65's risk exposure changed between July 2018 and May 2019.

\begin{table}[h!]
  \begin{center}
    \caption{Comparing risk evaluations from July '18 and May '19}
    \label{tab:table1}
    \begin{tabular}{l|l|l|l} % <-- Alignments: l=left, c=centre, r=right
    	\textbf{Risk} 	&	\textbf{July 2018} & \textbf{May 2019} & \textbf{Change}\\
      \hline
     Laws and regulations 				& High		& Low 		& Reduced\\
     Sourcing old components			& High		& Moderate	& Reduced \\
     Commodore logo						& Extreme	& Low  		& Reduced\\
     Commodore character set			& High		& Low  		& Reduced\\
     Commodore BASIC ROM				& High		& Low 		& Reduced \\
     Commodore kernel ROM				& High		& Low  		& Reduced\\
     Open-source business				& High		& Low  		& Reduced\\
    \end{tabular}
  \end{center}
\end{table}

\section{Discussion and Conclusion}
By comparing the risk evaluation results show in table \ref{tab:table1}, it can be seen that the MEGA65's risk exposure has been reduced in the 10 month period since the first risk evaluation was conducted. This reduction was caused by the MEGA65's team taking actions to mitigate the risks. These actions where, in part, caused by the advice and insight the MEGA65 team received in chapter \ref{Chapter5}. It is hoped that this outcome for the MEGA65 project helps justify the usefulness of this thesis work to retro-computing projects. This result also helps confirm the hypothesis stated at the start of this thesis. 

The outcomes for the MEGA65 project cannot be fully known at this stage, as the project has yet to be complete, it is also hard to determine exactly how effective the advice given in chapter \ref{Chapter5} was in reducing the risk exposure. Ideally if the MEGA65 project could be compared at a much later stage of production, the results could be of a higher quality. Also if the MEGA65 project could be compared with itself with and without the advice, then the resulting outcomes could provide more insight into the usefulness of this body of knowledge.

This thesis forms a body of knowledge on retro-computing projects the productization process, where there was no such body of knowledge before. It is hoped that this work will be continued and further research builds on it as there is many avenues of further research to pursue.